Results

### 2018-07-13 3-simus.Rmd
use a package

```{r, message=FALSE}
library(tidyverse)
library(glue)
library(bigsnpr)
```

@@

### PRS

@@

```{r}
PRS <- function(G, infos.chr, infos.pos, 
                pheno.all, covar.all, 
                ind.train, ind.test) {

  timing <- system.time({

    # GWAS
    gwas.train <- big_univLogReg(
      G, pheno.all[ind.train], ind.train = ind.train, 
      covar.train = covar.all[ind.train, , drop = FALSE], 
      ncores = NCORES
    )
    gwas.train.gc <- snp_gc(gwas.train)

  })[3]

  lapply(c(0.05, 0.2, 0.8), function(thr.r2) {

    # Clumping
    ind.keep <- snp_clumping(G, infos.chr = infos.chr,
                             ind.row = ind.test,
                             thr.r2 = thr.r2, 
                             S = abs(gwas.train.gc$score),
                             size = 500,
                             is.size.in.bp = TRUE,
                             infos.pos = infos.pos,
                             ncores = NCORES)

    # PRS
    thrs <- c(0, -log10(5e-8), exp(seq(log(0.1), log(100), length.out = 100)))
    lpS <- -predict(gwas.train.gc)
    prs <- snp_PRS(G, betas.keep = gwas.train.gc$estim[ind.keep],
                   ind.test = ind.test,
                   ind.keep = ind.keep,
                   lpS.keep = lpS[ind.keep], 
                   thr.list = thrs)

    ind.best <- which.max(apply(prs, 2, AUC, target = pheno.all[ind.test]))

    methods <- c("PRS-all", "PRS-stringent", "PRS-max")
    indices <- c(1:2, ind.best)

    lapply(1:3, function(i) {
      k <- indices[i]
      tibble(
        method = methods[i],
        pred = list(prs[, k]),
        thr.r2 = thr.r2,
        set = list(intersect(ind.keep, which(lpS > thrs[k]))),
        timing = timing
      )
    }) %>%
      bind_rows()
  }) %>%
    bind_rows()
}
```

### Logistic regression

```{r}
logit.CMSA <- function(G, pheno.all, covar.all, ind.train, ind.test, method,
                       alphas = c(1, 0.5, 0.05, 0.001)) {

  timing <- system.time({

    cmsa.logit <- big_spLogReg(X = G, y01.train = pheno.all[ind.train], 
                               ind.train = ind.train, 
                               covar.train = covar.all[ind.train, , drop = FALSE],
                               alphas = alphas, ncores = NCORES)

    preds <- predict(cmsa.logit, X = G, ind.row = ind.test, 
                     covar.row = covar.all[ind.test, , drop = FALSE])
  })[3]

  tibble(
    method   = method, 
    pred     = list(preds),
    timing   = timing,
    alpha = attr(cmsa.logit, "alpha"),
    set = list(which(rowSums(sapply(cmsa.logit, function(x) x$beta.X) != 0) > 0))
  )
}

```

### T-Trees

```{r}
bigstatsr:::assert_exist(TTree <- "../TTree-source/TTree")

ttrees <- function(TTree, file.base, pheno.all, ind.train, ind.test,
                   n.trees = 100) {

  file0.jdb  <- paste0(file.base, ".jdb")
  file0.bloc <- paste0(file.base, ".bloc")

  # Write jdb file with new pheno
  tmpfile <- tempfile()
  write(c(rep("", 5), pheno.all), ncolumns = 1, tmpfile)
  # https://stackoverflow.com/a/7846550/6103040
  system(sprintf("awk 'FNR==NR{a[NR]=$1;next}{$2=a[FNR]}1' %s %s > %s",
                 tmpfile, file0.jdb, file.jdb <- paste0(tmpfile, ".jdb")))
  # Write indices of learning and validation sets
  file.learn <- paste0(tmpfile, "_learn.txt")
  file.val   <- paste0(tmpfile, "_val.txt")
  cat(ind.train - 1, file = file.learn, sep = "\t")
  cat(ind.test - 1,  file = file.val,   sep = "\t")

  timing <- system.time(
    system(glue::glue(
      "{TTree}",
      " -j {file.jdb}",
      " -m 3",
      " -b {file0.bloc}",
      " -l {file.learn}",
      " -v {file.val}", 
      " -t {n.trees} -k 1000 -c 5 -n 2000",
      " -x -s"
    ))
  )[3]

  file.roc <- sprintf("%s_k1000_m3_t%d_ic5_nmin2000_0000.roc",
                      file.jdb, n.trees)
  file.vim <- sub("\\.roc$", ".vim", file.roc)
  preds <- read.table(file.roc, header = FALSE)

  tibble(
    method   = "T-Trees",
    pred     = list(preds[match(ind.test - 1, preds[[1]]), 2]),
    timing   = timing,
    set = list(which(read.table(file.vim)$V2 != 0))
  )
}
```

@@

@@

### Simulation of phenotypes

```{r}
get_pheno <- function(
  G,                                        ## matrix of genotypes
  h2,                                       ## heritability 
  M,                                        ## nbs of causal variants
  ind.possible = cols_along(G),             ## indices of possible causal variants
  effects.dist = c("gaussian", "laplace"),  ## distribution of effects 
  model = c("simple", "fancy"),             ## model for simulation
  K = 0.3                                   ## prevalence
) {

  effects.dist  <- match.arg(effects.dist)
  model <- match.arg(model)

  set <- sample(ind.possible, size = M)
  effects <- `if`(effects.dist == "gaussian", 
                  rnorm(M, sd = sqrt(h2 / M)),
                  rmutil::rlaplace(M, s = sqrt(h2 / (2*M))))

  if (model == "simple") {
    # only linear
    y.simu <- scale(G[, set]) %*% effects
  } else {
    sets <- split(1:M, sample(rep_len(1:3, M)))
    # linear
    ind1 <- sets[[1]]
    y.simu <- scale(G[, set[ind1]]) %*% effects[ind1]
    # recessive / dominant
    ind2 <- sets[[2]]
    y.simu <- y.simu + scale(G[, set[ind2]] > 0.5) %*% effects[ind2]
    # interactions
    ind3 <- matrix(sets[[3]], ncol = 2)
    y.simu <- y.simu + scale(G[, set[ind3[, 1]]] * G[, set[ind3[, 2]]]) %*% 
      effects[ind3[, 1]] * sqrt(2)
  }

  y.simu <- y.simu / sd(y.simu) * sqrt(h2)
  stopifnot(all.equal(drop(var(y.simu)), h2))
  y.simu <- y.simu + rnorm(nrow(G), sd = sqrt(1 - h2))
  list(pheno = as.integer(y.simu > qnorm(1 - K)), set = set, effects = effects)
}
```

@@

  # Simulate phenotypes
  simu_pheno <- get_pheno(
    G,    
    h2 = params[["par.h2"]], 
    M = as.integer(par.causal[1]), 
    ind.possible = `if`(par.causal[2] == "all", cols_along(G), ind.HLA),
    effects.dist = params[["par.dist"]], 
    model = params[["par.model"]]
  )
  pheno.all <- simu_pheno$pheno
  params[["set"]] <- list(simu_pheno$set)

@@

  # Get results from all methods
  res <- bind_rows(
    # PRS(G, CHR, POS, pheno.all, covar.all, ind.train, ind.test),
    logit.CMSA(G,  pheno.all, covar.all, ind.train, ind.test, "logit-simple"),
    logit.CMSA(G2, pheno.all, covar.all, ind.train, ind.test, "logit-triple"),
    ttrees(TTree, "backingfiles/ttrees", pheno.all, ind.train, ind.test, n.trees = 100)
  )
  params[["res"]] <- list(res)
  saveRDS(unnest(params, res, .drop = FALSE), file = res.file)
}
```

@@

  # Simulate phenotypes
  simu_pheno <- get_pheno(
    G,    
    h2 = params[["par.h2"]], 
    M = as.integer(par.causal[1]), 
    ind.possible = `if`(par.causal[2] == "all", cols_along(G), ind.HLA),
    effects.dist = params[["par.dist"]], 
    model = params[["par.model"]]
  )
  pheno.all <- simu_pheno$pheno
  params[["set"]] <- list(simu_pheno$set)

>>>>

```{r, message=FALSE}
library(tidyverse)
# devtools::install_github("privefl/bigsnpr")
library(bigsnpr)
# devtools::install_github("privefl/paper2-PRS/pkg.paper.PRS")
library(pkg.paper.PRS)  ## See code source for functions' code.
```

@@

@@

@@

TTree <- "../TTree-source/TTree"

@@

@@

  # Simulate phenotypes
  simu_pheno <- get_pheno(
    G,    
    h2 = params[["par.h2"]], 
    M = as.integer(par.causal[1]), 
    ind.possible = `if`(par.causal[2] == "all", cols_along(G), ind.HLA),
    effects.dist = params[["par.dist"]], 
    model = params[["par.model"]]
  )
  pheno.all <- simu_pheno$pheno
  params[["true_set"]] <- list(simu_pheno$set)

@@

  # Get results from all methods
  res <- bind_rows(
    # PRS(G, CHR, POS, pheno.all, covar.all, ind.train, ind.test),
    logit.CMSA(G,  pheno.all, covar.all, ind.train, ind.test, "logit-simple"),
    logit.CMSA(G2, pheno.all, covar.all, ind.train, ind.test, "logit-triple"),
    ttrees("../TTree-source/TTree", "backingfiles/ttrees", 
           pheno.all, ind.train, ind.test, n.trees = 100)
  )
  params[["res"]] <- list(res)
  saveRDS(unnest(params, res, .drop = FALSE), file = res.file)
}
```

@@

  # Simulate phenotypes
  simu_pheno <- get_pheno(
    G6,    
    h2 = params[["par.h2"]], 
    M = as.integer(par.causal[1]), 
    ind.possible = `if`(par.causal[2] == "all", cols_along(G6), ind.HLA6),
    effects.dist = params[["par.dist"]], 
    model = params[["par.model"]]
  )
  pheno.all <- simu_pheno$pheno
  params[["true_set"]] <- list(simu_pheno$set)

### 2018-07-13 4-results-simus.Rmd
save
use a package

```{r}
pAUC <- function(pred, target, p = 0.1) {
  val.min <- min(target)
  q <- quantile(pred[target == val.min], probs = 1 - p)
  ind <- (target != val.min) | (pred > q)
  tryCatch(bigstatsr::AUC(pred[ind], target[ind]) * p, 
           error = function(e) p^2 / 2)  # all same prediction
}

format_results <- function(results) {
  results %>%
    as_tibble() %>%
    mutate(
      par.causal = factor(map_chr(par.causal, ~paste(.x[1], .x[2], sep = " in ")),
                          levels = c("30 in HLA", paste(3 * 10^(1:3), "in all"))),
      AUC = map_dbl(eval, ~ bigstatsr::AUC(.x[, 1], .x[, 2])),
      pAUC = map_dbl(eval, ~ pAUC(.x[, 1], .x[, 2]))
    )
}
```

```{r}
boot <- function(x, n = 1e5, f = mean) {
  sd(replicate(n, f(sample(x, replace = TRUE))))
}
```

@@

```{r}
barplot_causal <- function(results, h2) {

>>>>

@@

```{r}
barplot_causal <- function(results) {
